---
title: "Brexit poll analysis"
output: html_notebook
---
Directions
There are 12 multi-part problems in this comprehensive assessment that review concepts from the entire course. The problems are split over 3 pages. Make sure you read the instructions carefully and run all pre-exercise code.

For numeric entry problems, you have 10 attempts to input the correct answer. For true/false problems, you have 2 attempts.

If you have questions, visit the "Brexit poll analysis" discussion forum that follows the assessment.

IMPORTANT: Some of these exercises use dslabs datasets that were added in a July 2019 update. Make sure your package is up to date with the command install.packages("dslabs").

Overview
In June 2016, the United Kingdom (UK) held a referendum to determine whether the country would "Remain" in the European Union (EU) or "Leave" the EU. This referendum is commonly known as Brexit. Although the media and others interpreted poll results as forecasting "Remain" (𝑝>0.5), the actual proportion that voted "Remain" was only 48.1% (𝑝=0.481) and the UK thus voted to leave the EU. Pollsters in the UK were criticized for overestimating support for "Remain". 

In this project, you will analyze real Brexit polling data to develop polling models to forecast Brexit results. You will write your own code in R and enter the answers on the edX platform.

Important definitions
Data Import
Import the brexit_polls polling data from the dslabs package and set options for the analysis:



```{r}
# suggested libraries and options
library(tidyverse)
options(digits = 3)

# load brexit_polls object
library(dslabs)
data(brexit_polls)
```

Final Brexit parameters
Define 𝑝=0.481 as the actual percent voting "Remain" on the Brexit referendum and 𝑑=2𝑝−1=−0.038 as the actual spread of the Brexit referendum with "Remain" defined as the positive outcome:

```{r}
p <- 0.481    # official proportion voting "Remain"
d <- 2*p-1    # official spread
```

Question 1: Expected value and standard error of a poll
0.0/6.0 points (graded)
The final proportion of voters choosing "Remain" was 𝑝=0.481. Consider a poll with a sample of 𝑁=1500 voters.

What is the expected total number of voters in the sample choosing "Remain"?

Explanation

You can calculate the expected number of "Remain" voters with the following code:

```{r}
  p <- 0.481
          N <- 1500
          N*p
```

What is the standard error of the total number of voters in the sample choosing "Remain"?


Explanation

You can calculate the standard error of the expected number of "Remain" voters with the following code:

```{r}
          sqrt(N*p*(1-p))

```

What is the expected value of  𝑋̂  , the proportion of "Remain" voters?

Explanation

The expected value of 𝑋̂  is 𝑝=0.481.



What is the standard error of  𝑋̂  , the proportion of "Remain" voters?

Explanation

You can calculate the standard error 𝑋̂  with the following code:

```{r}
          sqrt(p*(1-p)/N)

```

What is the expected value of  𝑑 , the spread between the proportion of "Remain" voters and "Leave" voters?

Explanation

Given the proportion 𝑝, the expected value of the spread is 2𝑝−1:

```{r}
          2*p-1

```

What is the standard error of  𝑑 , the spread between the proportion of "Remain" voters and "Leave" voters?

Explanation

The standard error of the spread is twice the standard error of 𝑋̂ :


```{r}
          2*sqrt(p*(1-p)/N)

```

Question 2: Actual Brexit poll estimates
0.0/4.0 points (graded)
Load and inspect the brexit_polls dataset from dslabs, which contains actual polling data for the 6 months before the Brexit vote. Raw proportions of voters preferring "Remain", "Leave", and "Undecided" are available (remain, leave, undecided) The spread is also available (spread), which is the difference in the raw proportion of voters choosing "Remain" and the raw proportion choosing "Leave".

Calculate x_hat for each poll, the estimate of the proportion of voters choosing "Remain" on the referendum day (𝑝=0.481), given the observed spread and the relationship 𝑑̂ =2𝑋̂ −1. Use mutate() to add a variable x_hat to the brexit_polls object by filling in the skeleton code below:

```{r}
brexit_polls <- brexit_polls %>%
        mutate(x_hat = __________)
```
What is the average of the observed spreads (spread)?

Answer code 

```{r}
brexit_polls <- brexit_polls %>%
          mutate(x_hat = (spread + 1)/2)
          mean(brexit_polls$spread)
        
```
What is the standard deviation of the observed spreads?

Answer code

```{r}
          sd(brexit_polls$spread)

```
What is the average of x_hat, the estimates of the parameter  


Answer code

```{r}
          mean(brexit_polls$x_hat)

```

What is the standard deviation of x_hat?

Answer code

```{r}
          sd(brexit_polls$x_hat)

```

Question 3: Confidence interval of a Brexit poll
3 points possible (graded)
Consider the first poll in brexit_polls, a YouGov poll run on the same day as the Brexit referendum:

```{r}
brexit_polls[1,]

```
Use qnorm() to compute the 95% confidence interval for 𝑋̂ .

What is the lower bound of the 95% confidence interval?

Answer code

```{r}
x_hat <- 0.52
N <- 4772
se_hat <- sqrt(x_hat*(1-x_hat)/N)
x_hat - qnorm(.975)*se_hat
```
What is the upper bound of the 95% confidence interval?


Answer code

```{r}
          x_hat + qnorm(.975)*se_hat

```

Does the 95% confidence interval predict a winner (does not cover  𝑝=0.5 )? Does the 95% confidence interval cover the true value of  𝑝  observed during the referendum?

Answer code

```{r}
!between(0.5, x_hat - qnorm(.975)*se_hat, x_hat + qnorm(.975)*se_hat)    # predicts winner
between(0.481, x_hat - qnorm(.975)*se_hat, x_hat + qnorm(.975)*se_hat)    # does not cover p

```

Brexit poll analysis - Part 2


Final Assessment due Jun 15, 2021 14:57 +03
This problem set is continued from the previous page. Make sure you have run the following code:

```{r}
# suggested libraries
library(tidyverse)

# load brexit_polls object and add x_hat column
library(dslabs)
data(brexit_polls)
brexit_polls <- brexit_polls %>%
    mutate(x_hat = (spread + 1)/2)

# final proportion voting "Remain"
p <- 0.481
```

Question 4: Confidence intervals for polls in June
0.0/4.0 points (graded)
Create the data frame june_polls containing only Brexit polls ending in June 2016 (enddate of "2016-06-01" and later). We will calculate confidence intervals for all polls and determine how many cover the true value of 𝑑.

First, use mutate() to calculate a plug-in estimate se_x_hat for the standard error of the estimate SE^[𝑋] for each poll given its sample size and value of 𝑋̂  (x_hat). Second, use mutate() to calculate an estimate for the standard error of the spread for each poll given the value of se_x_hat. Then, use mutate() to calculate upper and lower bounds for 95% confidence intervals of the spread. Last, add a column hit that indicates whether the confidence interval for each poll covers the correct spread 𝑑=−0.038.

How many polls are in june_polls?

Create june_polls and count the number of rows with the following code:

```{r}
june_polls <- brexit_polls %>%
              filter(enddate >= "2016-06-01")
    nrow(june_polls)
        
```
What proportion of polls have a confidence interval that covers the value 0?


Answer code

```{r}
 june_polls <- june_polls %>%
              mutate(se_x_hat = sqrt(x_hat*(1-x_hat)/samplesize),
                     se_spread = 2*se_x_hat,
                     ci_lower_spread = spread - qnorm(0.975)*se_spread,
                     ci_upper_spread = spread + qnorm(0.975)*se_spread)
          mean(june_polls$ci_lower_spread < 0 & june_polls$ci_upper_spread > 0)
        
```
What proportion of polls predict "Remain" (confidence interval entirely above 0)?


Answer code

```{r}
          mean(june_polls$ci_lower_spread > 0)

```

What proportion of polls have a confidence interval covering the true value of  

Answer code

```{r}
june_polls <- june_polls %>%
    mutate(hit = (2*p-1 > ci_lower_spread) & (2*p-1 < ci_upper_spread))
    mean(june_polls$hit)
```

Question 7: Combined spread across poll type
0.0/2.0 points (graded)
Calculate the confidence intervals of the spread combined across all polls in june_polls, grouping by poll type. Recall that to determine the standard error of the spread, you will need to double the standard error of the estimate.

Use this code (which determines the total sample size per poll type, gives each spread estimate a weight based on the poll's sample size, and adds an estimate of p from the combined spread) to begin your analysis:


```{r}
    
combined_by_type <- june_polls %>%
        group_by(poll_type) %>%
        summarize(N = sum(samplesize),
                  spread = sum(spread*samplesize)/N,
                  p_hat = (spread + 1)/2)
```
What is the lower bound of the 95% confidence interval for online voters?

Create june_polls and count the number of rows with the following code:

```{r}
  combined_by_type <- june_polls %>%
    group_by(poll_type) %>%
    summarize(N = sum(samplesize),
              spread = sum(spread*samplesize)/N,
              p_hat = (spread + 1)/2,
              se_spread = 2*sqrt(p_hat*(1-p_hat)/N),
              spread_lower = spread - qnorm(.975)*se_spread,
              spread_upper = spread + qnorm(.975)*se_spread)
combined_by_type %>%
    filter(poll_type == "Online") %>%
    pull(spread_lower)
```

What is the upper bound of the 95% confidence interval for online voters?

Answer code
```{r}
  combined_by_type %>%
    filter(poll_type == "Online") %>%
    pull(spread_upper)
```

Brexit poll analysis - Part 3


Final Assessment due Jun 15, 2021 14:57 +03
This problem set is continued from the previous page. Make sure you have run the following code:


```{r}
# suggested libraries
library(tidyverse)

# load brexit_polls object and add x_hat column
library(dslabs)
data(brexit_polls)
brexit_polls <- brexit_polls %>%
    mutate(x_hat = (spread + 1)/2)

# final proportion voting "Remain"
p <- 0.481
```
Question 9: Chi-squared p-value
0.0/2.0 points (graded)
Define brexit_hit, with the following code, which computes the confidence intervals for all Brexit polls in 2016 and then calculates whether the confidence interval covers the actual value of the spread 𝑑=−0.038:


```{r}
brexit_hit <- brexit_polls %>%
  mutate(p_hat = (spread + 1)/2,
         se_spread = 2*sqrt(p_hat*(1-p_hat)/samplesize),
         spread_lower = spread - qnorm(.975)*se_spread,
         spread_upper = spread + qnorm(.975)*se_spread,
         hit = spread_lower < -0.038 & spread_upper > -0.038) %>%
  select(poll_type, hit)
```
Use brexit_hit to make a two-by-two table of poll type and hit status. Then use the chisq.test() function to perform a chi-squared test to determine whether the difference in hit rate is significant.

What is the p-value of the chi-squared test comparing the hit rate of online and telephone polls?

Answer code

```{r}
          brexit_chisq <- table(brexit_hit$poll_type, brexit_hit$hit)
chisq.test(brexit_chisq)$p.value
        
```
Determine which poll type has a higher probability of producing a confidence interval that covers the correct value of the spread. Also determine whether this difference is statistically significant at a p-value cutoff of 0.05. Which of the following is true?


Answer code

```{r}
# online > telephone
hit_rate <- brexit_hit %>%
    group_by(poll_type) %>%
    summarize(avg = mean(hit))
hit_rate$avg[hit_rate$poll_type == "Online"] > hit_rate$avg[hit_rate$poll_type == "Telephone"]

# statistically significant
chisq.test(brexit_chisq)$p.value < 0.05
```

Question 10: Odds ratio of online and telephone poll hit rate
0.0/3.0 points (graded)
Use the two-by-two table constructed in the previous exercise to calculate the odds ratio between the hit rate of online and telephone polls to determine the magnitude of the difference in performance between the poll types.

Calculate the odds that an online poll generates a confidence interval that covers the actual value of the spread.

Answer code

```{r}
# from previous question
brexit_chisq <- table(brexit_hit$poll_type, brexit_hit$hit)

# convert to data frame
chisq_df <- as.data.frame(brexit_chisq)

online_true <- chisq_df$Freq[chisq_df$Var1 == "Online" & chisq_df$Var2 == "TRUE"]
online_false <- chisq_df$Freq[chisq_df$Var1 == "Online" & chisq_df$Var2 == "FALSE"]

online_odds <- online_true/online_false
online_odds
        
```
Calculate the odds that a telephone poll generates a confidence interval that covers the actual value of the spread.


Answer code

```{r}
phone_true <- chisq_df$Freq[chisq_df$Var1 == "Telephone" & chisq_df$Var2 == "TRUE"]
phone_false <- chisq_df$Freq[chisq_df$Var1 == "Telephone" & chisq_df$Var2 == "FALSE"]

phone_odds <- phone_true/phone_false
phone_odds
        
```
Calculate the odds ratio to determine how many times larger the odds are for online polls to hit versus telephone polls.

Answer code

```{r}
online_odds/phone_odds

```
Question 11: Plotting spread over time
1/1 point (graded)
Use brexit_polls to make a plot of the spread (spread) over time (enddate) colored by poll type (poll_type). Use geom_smooth() with method = "loess" to plot smooth curves with a span of 0.4. Include the individual data points colored by poll type. Add a horizontal line indicating the final value of 𝑑=−.038.

Which of the following plots is correct?


Answer code

```{r}
brexit_polls %>%
              ggplot(aes(enddate, spread, color = poll_type)) +
              geom_smooth(method = "loess", span = 0.4) +
              geom_point() +
              geom_hline(aes(yintercept = -.038))
```
Question 12: Plotting raw percentages over time
0.0/2.5 points (graded)
Use the following code to create the object brexit_long, which has a column vote containing the three possible votes on a Brexit poll ("remain", "leave", "undecided") and a column proportion containing the raw proportion choosing that vote option on the given poll:

```{r}
brexit_long <- brexit_polls %>%
    gather(vote, proportion, "remain":"undecided") %>%
    mutate(vote = factor(vote))
```

Make a graph of proportion over time colored by vote. Add a smooth trendline with geom_smooth() and method = "loess" with a span of 0.3.

Which of the following are TRUE?