---
title: "Assessment 2.1: Introduction to InferenceExercise 1. Sample average
Write function called take_sample that takes the proportion of Democrats  and the sample size  as arguments and returns the sample average of Democrats (1) and Republicans (0).

Calculate the sample average if the proportion of Democrats equals 0.45 and the sample size is 100."
output: html_notebook
---

```{r}
# Write a function called `take_sample` that takes `p` and `N` as arguements and returns the average value of a randomly sampled population.
take_sample <- function(p, N){
    X <- sample(c(0,1), size = N, replace = TRUE, prob = c(1 - p, p))
    mean(X)
}

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Define `p` as the proportion of Democrats in the population being polled
p <- 0.45

# Define `N` as the number of people polled
N <- 100

# Call the `take_sample` function to determine the sample average of `N` randomly selected people from a population containing a proportion of Democrats equal to `p`. Print this value to the console.
take_sample(p,N)

```
Exercise 2. Distribution of errors - 1
Assume the proportion of Democrats in the population  equals 0.45 and that your sample size  is 100 polled voters. The take_sample function you defined previously generates our estimate, 
.

Replicate the random sampling 10,000 times and calculate 
 for each random sample. Save these differences as a vector called errors. Find the average of errors and plot a histogram of the distribution.
 
```{r}
# Define `p` as the proportion of Democrats in the population being polled
p <- 0.45

# Define `N` as the number of people polled
N <- 100

# The variable `B` specifies the number of times we want the sample to be replicated
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Create an objected called `errors` that replicates subtracting the result of the `take_sample` function from `p` for `B` replications
errors <- replicate(B, p - take_sample(p, N))

# Calculate the mean of the errors. Print this value to the console.
mean(errors)


```
 Exercise 4. Average size of error
The error 
 is a random variable. In practice, the error is not observed because we do not know the actual proportion of Democratic voters, . However, we can describe the size of the error by constructing a simulation.

What is the average size of the error if we define the size by taking the absolute value  p-x
 ?

```{r}
# Define `p` as the proportion of Democrats in the population being polled
p <- 0.45

# Define `N` as the number of people polled
N <- 100

# The variable `B` specifies the number of times we want the sample to be replicated
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# We generated `errors` by subtracting the estimate from the actual proportion of Democratic voters
errors <- replicate(B, p - take_sample(p, N))

# Calculate the mean of the absolute value of each simulated error. Print this value to the console.
mean(abs(errors))

```
Exercise 5. Standard deviation of the spread
The standard error is related to the typical size of the error we make when predicting. We say size because, as we just saw, the errors are centered around 0. In that sense, the typical error is 0. For mathematical reasons related to the central limit theorem, we actually use the standard deviation of errors rather than the average of the absolute values.

As we have discussed, the standard error is the square root of the average squared distance 
. The standard deviation is defined as the square root of the distance squared.

Calculate the standard deviation of the spread.

```{r}
# Define `p` as the proportion of Democrats in the population being polled
p <- 0.45

# Define `N` as the number of people polled
N <- 100

# The variable `B` specifies the number of times we want the sample to be replicated
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# We generated `errors` by subtracting the estimate from the actual proportion of Democratic voters
errors <- replicate(B, p - take_sample(p, N))

# Calculate the standard deviation of `errors`
sqrt(mean(errors^2))

```
Exercise 6. Estimating the standard error
The theory we just learned tells us what this standard deviation is going to be because it is the standard error of 
.

Estimate the standard error given an expected value of 0.45 and a sample size of 100.

```{r}
# Define `p` as the expected value equal to 0.45
p <- 0.45

# Define `N` as the sample size
N <- 100

# Calculate the standard error
sqrt(p*(1-p)/N)

```

Exercise 7. Standard error of the estimate
In practice, we don't know , so we construct an estimate of the theoretical prediction based by plugging in 
 for . Calculate the standard error of the estimate:

```{r}
# Define `p` as a proportion of Democratic voters to simulate
p <- 0.45

# Define `N` as the sample size
N <- 100

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Define `X` as a random sample of `N` voters with a probability of picking a Democrat ('1') equal to `p`
X <- sample(c(0,1), size=N, replace=TRUE, prob=c(1-p, p))

# Define `X_bar` as the average sampled proportion
X_bar <- mean(X)

# Calculate the standard error of the estimate. Print the result to the console.
sqrt(X_bar*(1-X_bar)/N)

```
Exercise 11. Plotting the errors
Make a qq-plot of the errors you generated previously to see if they follow a normal distribution.

```{r}
# Define `p` as the proportion of Democrats in the population being polled
p <- 0.45

# Define `N` as the number of people polled
N <- 100

# The variable `B` specifies the number of times we want the sample to be replicated
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Generate `errors` by subtracting the estimate from the actual proportion of Democratic voters
errors <- replicate(B, p - take_sample(p, N))

# Generate a qq-plot of `errors` with a qq-line showing a normal distribution
qqnorm(errors)
qqline(errors)
```
Exercise 12. Estimating the probability of a specific value of X-bar
If  and , use the central limit theorem to estimate the probability that 
.

```{r}
# Define `p` as the proportion of Democrats in the population being polled
p <- 0.45

# Define `N` as the number of people polled
N <- 100

# Calculate the probability that the estimated proportion of Democrats in the population is greater than 0.5. Print this value to the console.
1 - pnorm(0.5, p, sqrt(p*(1-p)/N))
```
Exercise 13. Estimating the probability of a specific error size
Assume you are in a practical situation and you don't know . Take a sample of size  and obtain a sample average of 
.

What is the CLT approximation for the probability that your error size is equal or larger than 0.01?

```{r}
# Define `N` as the number of people polled
N <-100

# Define `X_hat` as the sample average
X_hat <- 0.51

# Define `se_hat` as the standard error of the sample average
se_hat <- sqrt(X_hat*(1-X_hat)/N)

# Calculate the probability that the error is 0.01 or larger
1 - pnorm(.01, 0, se_hat) + pnorm(-0.01, 0, se_hat)

```

