---
title: "Assessment 4.1: Statistical Models"
output: html_notebook
---

Exercise 1 - Heights Revisited
We have been using urn models to motivate the use of probability models. However, most data science applications are not related to data obtained from urns. More common are data that come from individuals. Probability plays a role because the data come from a random sample. The random sample is taken from a population and the urn serves as an analogy for the population.

Let's revisit the heights dataset. For now, consider x to be the heights of all males in the data set. Mathematically speaking, x is our population. Using the urn analogy, we have an urn with the values of x in it.

What are the population average and standard deviation of our population?


```{r}
# Load the 'dslabs' package and data contained in 'heights'
library(dslabs)
data(heights)

# Make a vector of heights from all males in the population
x <- heights %>% filter(sex == "Male") %>%
  .$height

# Calculate the population average. Print this value to the console.
mean(x)

# Calculate the population standard deviation. Print this value to the console.
sd(x)
```
Exercise 2 - Sample the population of heights
Call the population average computed above  and the standard deviation . Now take a sample of size 50, with replacement, and construct an estimate for  and .

```{r}
# The vector of all male heights in our population `x` has already been loaded for you. You can examine the first six elements using `head`.
head(x)

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Define `N` as the number of people measured
N <- 50

# Define `X` as a random sample from our population `x`
X <- sample(x, N, replace = TRUE)

# Calculate the sample average. Print this value to the console.
mean(X)

# Calculate the sample standard deviation. Print this value to the console.
sd(X)
```
Exercise 4 - Confidence Interval Calculation
We will use 
 as our estimate of the heights in the population from our sample size . We know from previous exercises that the standard estimate of our error 
 is .

Construct a 95% confidence interval for .

```{r}
# The vector of all male heights in our population `x` has already been loaded for you. You can examine the first six elements using `head`.
head(x)

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Define `N` as the number of people measured
N <- 50

# Define `X` as a random sample from our population `x`
X <- sample(x, N, replace = TRUE)

# Define `se` as the standard error of the estimate. Print this value to the console.
se <- sd(X)/sqrt(N)
se

# Construct a 95% confidence interval for the population average based on our sample. Save the lower and then the upper confidence interval to a variable called `ci`.
ci <- c(mean(X) - qnorm(0.975)*se, mean(X) + qnorm(0.975)*se)

```
Exercise 5 - Monte Carlo Simulation for Heights
Now run a Monte Carlo simulation in which you compute 10,000 confidence intervals as you have just done. What proportion of these intervals include ?

```{r}
# Define `mu` as the population average
mu <- mean(x)

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Define `N` as the number of people measured
N <- 50

# Define `B` as the number of times to run the model
B <- 10000

# Define an object `res` that contains a logical vector for simulated intervals that contain mu
res <- replicate(B, {
  X <- sample(x, N, replace=TRUE)
  interval <- mean(X) + c(-1,1)*qnorm(0.975)*sd(X)/sqrt(N)
  between(mu, interval[1], interval[2])
})

# Calculate the proportion of results in `res` that include mu. Print this value to the console.
mean(res)
```

Exercise 6 - Visualizing Polling Bias
In this section, we used visualization to motivate the presence of pollster bias in election polls. Here we will examine that bias more rigorously. Lets consider two pollsters that conducted daily polls and look at national polls for the month before the election.

Is there a poll bias? Make a plot of the spreads for each poll.
```{r}
# These lines of code filter for the polls we want and calculate the spreads
polls <- polls_us_election_2016 %>% 
  filter(pollster %in% c("Rasmussen Reports/Pulse Opinion Research","The Times-Picayune/Lucid") &
           enddate >= "2016-10-15" &
           state == "U.S.") %>% 
  mutate(spread = rawpoll_clinton/100 - rawpoll_trump/100) 

# Make a boxplot with points of the spread for each pollster
polls %>% ggplot(aes(pollster, spread)) + 
  geom_boxplot() + 
  geom_point()
```
Exercise 13 - Compute the Estimates
The answer to the previous question depends on 
 and 
, which we don't know. We learned that we can estimate these values using the sample standard deviation.

Compute the estimates of 
 and 
 
```{r}
# The `polls` data have already been loaded for you. Use the `head` function to examine them.
head(polls)

# Create an object called `sigma` that contains a column for `pollster` and a column for `s`, the standard deviation of the spread
sigma <- polls %>% group_by(pollster) %>%
  summarize(s = sd(spread))

# Print the contents of sigma to the console
sigma
```
 
Exercise 15 - Calculate the 95% Confidence Interval of the Spreads
We have constructed a random variable that has expected value 
, the pollster bias difference. If our model holds, then this random variable has an approximately normal distribution. The standard error of this random variable depends on 
 and 
, but we can use the sample standard deviations we computed earlier. We have everything we need to answer our initial question: is 
 different from 0?

Construct a 95% confidence interval for the difference 
 and 
. Does this interval contain zero?

```{r}
# The `polls` data have already been loaded for you. Use the `head` function to examine them.
head(polls)

# Create an object called `res` that summarizes the average, standard deviation, and number of polls for the two pollsters.
res <- polls %>% group_by(pollster) %>% 
  summarize(avg = mean(spread), s = sd(spread), N = n()) 

# Store the difference between the larger average and the smaller in a variable called `estimate`. Print this value to the console.
estimate <- res$avg[2] - res$avg[1]
estimate

# Store the standard error of the estimates as a variable called `se_hat`. Print this value to the console.
se_hat <- sqrt(res$s[2]^2/res$N[2] + res$s[1]^2/res$N[1])
se_hat

# Calculate the 95% confidence interval of the spreads. Save the lower and then the upper confidence interval to a variable called `ci`.
ci <- c(estimate - qnorm(0.975)*se_hat, estimate + qnorm(0.975)*se_hat)
```
Exercise 16 - Calculate the P-value
The confidence interval tells us there is relatively strong pollster effect resulting in a difference of about 5%. Random variability does not seem to explain it.

Compute a p-value to relay the fact that chance does not explain the observed pollster effect.

```{r}
# We made an object `res` to summarize the average, standard deviation, and number of polls for the two pollsters.
res <- polls %>% group_by(pollster) %>% 
  summarize(avg = mean(spread), s = sd(spread), N = n()) 

# The variables `estimate` and `se_hat` contain the spread estimates and standard error, respectively.
estimate <- res$avg[2] - res$avg[1]
se_hat <- sqrt(res$s[2]^2/res$N[2] + res$s[1]^2/res$N[1])

# Calculate the p-value
2*(1 - pnorm(estimate/se_hat, 0, 1))
```
Exercise 17 - Comparing Within-Poll and Between-Poll Variability
We compute statistic called the t-statistic by dividing our estimate of 
 by its estimated standard error:

 
Later we learn will learn of another approximation for the distribution of this statistic for values of 
 and 
 that aren't large enough for the CLT.

Note that our data has more than two pollsters. We can also test for pollster effect using all pollsters, not just two. The idea is to compare the variability across polls to variability within polls. We can construct statistics to test for effects and approximate their distribution. The area of statistics that does this is called Analysis of Variance or ANOVA. We do not cover it here, but ANOVA provides a very useful set of tools to answer questions such as: is there a pollster effect?

Compute the average and standard deviation for each pollster and examine the variability across the averages and how it compares to the variability within the pollsters, summarized by the standard deviation.

```{r}
# Execute the following lines of code to filter the polling data and calculate the spread
polls <- polls_us_election_2016 %>% 
  filter(enddate >= "2016-10-15" &
           state == "U.S.") %>%
  group_by(pollster) %>%
  filter(n() >= 5) %>% 
  mutate(spread = rawpoll_clinton/100 - rawpoll_trump/100) %>%
  ungroup()

# Create an object called `var` that contains columns for the pollster, mean spread, and standard deviation. Print the contents of this object to the console.

var <- polls %>% group_by(pollster) %>%
  summarize(avg = mean(spread), s = sd(spread))
var

```

