---
title: "DataCamp Assessment 1.1: Parameters and Estimates"
output: html_notebook
---
Exercise 5. se versus p
Write a line of code that calculates the standard error se of a sample average when you poll 25 people in the population. Generate a sequence of 100 proportions of Democrats p that vary from 0 (no Democrats) to 1 (all Democrats).

Plot se versus p for the 100 different proportions.

```{r}
# `N` represents the number of people polled
N <- 25

# Create a variable `p` that contains 100 proportions ranging from 0 to 1 using the `seq` function
p <- seq(0, 1, length = 100)

# Create a variable `se` that contains the standard error of each sample average
se <- sqrt(p*(1-p)/N)

# Plot `p` on the x-axis and `se` on the y-axis
plot(p, se)
```
Exercise 6. Multiple plots of se versus p
Using the same code as in the previous exercise, create a for-loop that generates three plots of p versus se when the sample sizes equal N=25 ,N=100 , and . N = 1000

```{r}
# The vector `p` contains 100 proportions of Democrats ranging from 0 to 1 using the `seq` function
p <- seq(0, 1, length = 100)

# The vector `sample_sizes` contains the three sample sizes
sample_sizes <- c(25, 100, 1000)

# Write a for-loop that calculates the standard error `se` for every value of `p` for each of the three samples sizes `N` in the vector `sample_sizes`. Plot the three graphs, using the `ylim` argument to standardize the y-axis across all three plots.
for(N in sample_sizes){
  se <- sqrt(p*(1-p)/N)
  plot(p, se, ylim = c(0,0.5/sqrt(25)))
}
```

Exercise 9. Standard error of the spread
Say the actual proportion of Democratic voters is . In this case, the Republican party is winning by a relatively large margin of , or a 10% margin of victory. What is the standard error of the spread 
 in this case?
 
```{r}
# `N` represents the number of people polled
N <- 25

# `p` represents the proportion of Democratic voters
p <- 0.45

# Calculate the standard error of the spread. Print this value to the console.
2*sqrt(p*(1-p)/N)


```
 
