---
title: "Introduction to Assessment: The Big Short"
output: html_notebook
---

Introduction
These exercises, available to verified learners only, review and assess the following concepts:

Expected value and standard error of a single draw of a random variable
Expected value and standard error of the sum of draws of a random variable
Monte Carlo simulation of the sum of draws of a random variable
The Central Limit Theorem approximation of the sum of draws of a random variable
Using z-scores to calculate values related to the normal distribution and normal random variables
Calculating interest/premium rates to minimize chance of losing money
Determining a number of loans/policies required to profit
Simulating the effects of a change in event probability

Setup and libraries
Run the code below to set up your environment and load the libraries you will need for the following exercises:

```{r}
library(tidyverse)
library(dslabs)
```
Background
In the motivating example The Big Short, we discussed how discrete and continuous probability concepts relate to bank loans and interest rates. Similar business problems are faced by the insurance industry. 

Just as banks must decide how much to charge as interest on loans based on estimates of loan defaults, insurance companies must decide how much to charge as premiums for policies given estimates of the probability that an individual will collect on that policy. 

We will use data from 2015 US Period Life Tables External link. Here is the code you will need to load and examine the data from dslabs:

```{r}
data(death_prob)
head(death_prob)
```
Questions 1 and 2: Insurance rates, part 1

Use the information below as you answer this 6-part question.

An insurance company offers a one-year term life insurance policy that pays $150,000 in the event of death within one year. The premium (annual cost) for this policy for a 50 year old female is $1,150. Suppose that in the event of a claim, the company forfeits the premium and loses a total of $150,000, and if there is no claim the company gains the premium amount of $1,150. The company plans to sell 1,000 policies to this demographic.

Question 1a
0.0/1.0 point (graded)
The death_prob data frame from the dslabs package contains information about the estimated probability of death within 1 year (prob) for different ages and sexes.

Use death_prob to determine the death probability of a 50 year old female, p.


Explanation

The probability can be calculated using the following code:

```{r}

 p <- death_prob %>%
  filter(sex == "Female" & age == "50") %>%
  pull(prob)
p
```
Question 1b
1/1 point (graded)
The loss in the event of the policy holder's death is -$150,000 and the gain if the policy holder remains alive is the premium $1,150.

What is the expected value of the company's net profit on one policy for a 50 year old female?

Explanation

The expected value can be calculated using the following code:

```{r}
a <- -150000
b <- 1150

mu <- a*p + b*(1-p)
mu
```
Question 1c
1/1 point (graded)
Calculate the standard error of the profit on one policy for a 50 year old female.


```{r}
sigma <- abs(b-a) * sqrt(p*(1-p))
sigma
        
```

Question 1d
0.0/1.0 point (graded)
What is the expected value of the company's profit over all 1,000 policies for 50 year old females?


```{r}
n <- 1000
n*mu
```

Question 1e
0.0/1.0 point (graded)
What is the standard error of the sum of the expected value over all 1,000 policies for 50 year old females?

```{r}
          sqrt(n) * sigma

```

Question 1f
0.0/1.0 point (graded)
Use the Central Limit Theorem to calculate the probability that the insurance company loses money on this set of 1,000 policies.


```{r}
          pnorm(0, n*mu, sqrt(n)*sigma)

```

50 year old males have a different probability of death than 50 year old females. We will calculate a profitable premium for 50 year old males in the following four-part question.

Question 2a
0.0/1.0 point (graded)
Use death_prob to determine the probability of death within one year for a 50 year old male.

```{r}
 p_male <- death_prob %>%
  filter(sex == "Male" & age == "50") %>%
  pull(prob)
p_male
```

Question 2b
0.0/1.0 point (graded)
Suppose the company wants its expected profits from 1,000 50 year old males with $150,000 life insurance policies to be $700,000. Use the formula for expected value of the sum of draws with the following values and solve for the premium ğ‘:

E[ğ‘†]=ğœ‡ğ‘†=700000
ğ‘›=1000
ğ‘=death probability of age 50 male
ğ‘=150000
ğ‘=premium
What premium should be charged?

Explanation

Start with the formula for expected value of the sum and solve for ğ‘:

ğœ‡ğ‘†=ğ‘›(ğ‘ğ‘+ğ‘(1âˆ’ğ‘))
ğœ‡ğ‘†ğ‘›=ğ‘ğ‘+ğ‘(1âˆ’ğ‘)
ğœ‡ğ‘†ğ‘›âˆ’ğ‘ğ‘1âˆ’ğ‘=ğ‘
Then substitute known values for ğœ‡ğ‘†,ğ‘›,ğ‘,ğ‘ using the following code:


```{r}
p <- p_male
mu_sum <- 700000
n <- 1000
a <- -150000

b <- (mu_sum/n-a*p)/(1-p)
b
```
Question 2c
0.0/1.0 point (graded)
Using the new 50 year old male premium rate, calculate the standard error of the sum of 1,000 premiums.

```{r}
sigma_sum <- sqrt(n)*abs(b-a)*sqrt(p*(1-p))
sigma_sum
        
```
Question 2d
0.0/1.0 point (graded)
What is the probability of losing money on a series of 1,000 policies to 50 year old males?
Use the Central Limit Theorem.

Explanation

The probability can be calculated using the following code:

```{r}
          pnorm(0, mu_sum, sigma_sum)

```

Questions 3 and 4: insurance rates, part 2


Life insurance rates are calculated using mortality statistics from the recent past. They are priced such that companies are almost assured to profit as long as the probability of death remains similar. If an event occurs that changes the probability of death in a given age group, the company risks significant losses.

In this 6-part question, we'll look at a scenario in which a lethal pandemic disease increases the probability of death within 1 year for a 50 year old to .015. Unable to predict the outbreak, the company has sold 1,000 $150,000 life insurance policies for $1,150.

Question 3a
0.0/1.0 point (graded)
What is the expected value of the company's profits over 1,000 policies?

Explanation

The expected value can be calculated using the following code:

```{r}
 p <- .015    # probability of claim
a <- -150000    # loss per claim
b <- 1150    # premium - profit when no claim
n <- 1000

exp_val <- n*(a*p + b*(1-p))
exp_val
```

Question 3b
0.0/1.0 point (graded)
What is the standard error of the expected value of the company's profits over 1,000 policies?

Explanation

The standard error can be calculated using the following code:

```{r}
se <- sqrt(n) * abs(b-a) * sqrt(p*(1-p))
se
        
```

Question 3c
1/1 point (graded)
What is the probability of the company losing money?


Explanation

The probability can be calculated using the following code:

```{r}
          pnorm(0, exp_val, se)

```

Question 3d
1/1 point (graded)
Suppose the company can afford to sustain one-time losses of $1 million, but larger losses will force it to go out of business.

What is the probability of losing more than $1 million?
Explanation

The probability can be calculated using the following code:

```{r}
          pnorm(-1*10^6, exp_val, se)

```
Question 3e
0.0/1.0 point (graded)
Investigate death probabilities p <- seq(.01, .03, .001).

What is the lowest death probability for which the chance of losing money exceeds 90%?
  unanswered 

```{r}
 p <- seq(.01, .03, .001)
a <- -150000    # loss per claim
b <- 1150    # premium - profit when no claim
n <- 1000

p_lose_money <- sapply(p, function(p){
  exp_val <- n*(a*p + b*(1-p))
  se <- sqrt(n) * abs(b-a) * sqrt(p*(1-p))
  pnorm(0, exp_val, se)
})

data.frame(p, p_lose_money) %>%
  filter(p_lose_money > 0.9) %>%
  pull(p) %>%
  min()
```
Question 3f
1/1 point (graded)
Investigate death probabilities p <- seq(.01, .03, .0025).

What is the lowest death probability for which the chance of losing over $1 million exceeds 90%?

```{r}
 p_lose_million <- sapply(p, function(p){
  exp_val <- n*(a*p + b*(1-p))
  se <- sqrt(n) * abs(b-a) * sqrt(p*(1-p))
  pnorm(-1*10^6, exp_val, se)
})

data.frame(p, p_lose_million) %>%
  filter(p_lose_million > 0.9) %>%
  pull(p) %>%
  min()
        
```
Question 4a
1/1 point (graded)
Define a sampling model for simulating the total profit over 1,000 loans with probability of claim p_loss = .015, loss of -$150,000 on a claim, and profit of $1,150 when there is no claim. Set the seed to 25, then run the model once.

(IMPORTANT! If you use R 3.6 or later, you will need to use the command set.seed(x, sample.kind = "Rounding") instead of set.seed(x). Your R version will be printed at the top of the Console window when you start RStudio.)

What is the reported profit (or loss) in millions (that is, divided by 106)?

```{r}
 set.seed(25)

p <- .015
loss <- -150000
profit <- 1150
n <- 1000

outcomes <- sample(c(loss, profit), n, prob = c(p, 1-p), replace = TRUE)
sum(outcomes)/10^6
        
```

Question 4b
1/1 point (graded)
Set the seed to 27, then run a Monte Carlo simulation of your sampling model with 10,000 replicates to simulate the range of profits/losses over 1,000 loans.

(IMPORTANT! If you use R 3.6 or later, you will need to use the command set.seed(x, sample.kind = "Rounding") instead of set.seed(x). Your R version will be printed at the top of the Console window when you start RStudio.)

What is the observed probability of losing $1 million or more?

```{r}
set.seed(27)
B <- 10000

profits <- replicate(B, {
  outcomes <- sample(c(loss, profit), n, prob = c(p, 1-p), replace = TRUE)
  sum(outcomes)/10^6
})

mean(profits < -1)
        
```

Questions 5 and 6: Insurance rates, part 3

Question 5, which has 4 parts, continues the pandemic scenario from Questions 3 and 4.

Suppose that there is a massive demand for life insurance due to the pandemic, and the company wants to find a premium cost for which the probability of losing money is under 5%, assuming the death rate stays stable at ğ‘=0.015.

Question 5a
0.0/1.0 point (graded)
Calculate the premium required for a 5% chance of losing money given ğ‘›=1000 loans, probability of death ğ‘=0.015, and loss per claim ğ‘™=âˆ’150000. Save this premium as x for use in further questions.

```{r}
p <- .015
n <- 1000
l <- -150000
z <- qnorm(.05)
x <- -l*( n*p - z*sqrt(n*p*(1-p)))/ ( n*(1-p) + z*sqrt(n*p*(1-p)))
x
        
```

Question 5b
1/1 point (graded)
What is the expected profit per policy at this rate?

```{r}
          l*p + x*(1-p)

```
Question 5c
0.0/1.0 point (graded)
What is the expected profit over 1,000 policies?

```{r}
 mu <- n*(l*p + x*(1-p))
mu
        
```
Question 5d
0.0/1.0 point (graded)
Run a Monte Carlo simulation with B=10000to determine the probability of losing money on 1,000 policies given the new premium x, loss on a claim of $150,000, and probability of claim ğ‘=.015. Set the seed to 28 before running your simulation.

(IMPORTANT! If you use R 3.6 or later, you will need to use the command set.seed(x, sample.kind = "Rounding") instead of set.seed(x). Your R version will be printed at the top of the Console window when you start RStudio.)

What is the probability of losing money here?

```{r}
set.seed(28)
B <- 10000
profit <- replicate(B, {
  draws <- sample(c(x, l), n,
                  prob=c(1-p, p), replace = TRUE)
  sum(draws)
})

mean(profit < 0)
        
```
The company cannot predict whether the pandemic death rate will stay stable. Set the seed to 29, then write a Monte Carlo simulation that for each of ğµ=10000 iterations:

randomly changes ğ‘ by adding a value between -0.01 and 0.01 with sample(seq(-0.01, 0.01, length = 100), 1)
uses the new random ğ‘ to generate a sample of ğ‘›=1,000 policies with premium x and loss per claim ğ‘™=âˆ’150000
returns the profit over ğ‘› policies (sum of random variable)
(IMPORTANT! If you use R 3.6 or later, you will need to use the command set.seed(x, sample.kind = "Rounding") instead of set.seed(x). Your R version will be printed at the top of the Console window when you start RStudio.)

The outcome should be a vector of ğµ total profits. Use the results of the Monte Carlo simulation to answer the following three questions.

(Hint: Use the process from lecture for modeling a situation for loans that changes the probability of default for all borrowers simultaneously.)

Question 6a
0.0/1.0 point (graded)
What is the expected value over 1,000 policies?

```{r}
 # n, p, l and x as defined in the problem information
set.seed(29)    # in R 3.6, set.seed(29, sample.kind="Rounding")
  
profit <- replicate(B, {
    new_p <- p + sample(seq(-0.01, 0.01, length = 100), 1)
    draws <- sample(c(x, l), n, 
                    prob=c(1-new_p, new_p), replace = TRUE) 
    sum(draws)
})
```

This code gives the expected value for the profit:

```{r}
          mean(profit)

```
Question 6b
0.0/1.0 point (graded)
What is the probability of losing money?

```{r}
          mean(profit < 0)

```

Question 6c
0.0/1.0 point (graded)
What is the probability of losing more than $1 million?


```{r}
          mean(profit < -1*10^6)

```

