---
title: "Code: Function for birthday problem Monte Carlo simulations"
output: html_notebook
---

Code: Function for birthday problem Monte Carlo simulations
Note that the function body of compute_prob() is the code that we wrote in the previous video. If we write this code as a function, we can use sapply() to apply this function to several values of n.

```{r}
# function to calculate probability of shared bdays across n people
compute_prob <- function(n, B = 10000) {
	same_day <- replicate(B, {
    	bdays <- sample(1:365, n, replace = TRUE)
        any(duplicated(bdays))
    })
    mean(same_day)
}

n <- seq(1, 60)
```

Code: Element-wise operation over vectors and sapply


```{r}
x <- 1:10
sqrt(x)    # sqrt operates on each element of the vector

y <- 1:10
x*y    # * operates element-wise on both vectors

compute_prob(n)    # does not iterate over the vector n without sapply

x <- 1:10
sapply(x, sqrt)    # this is equivalent to sqrt(x)

prob <- sapply(n, compute_prob)    # element-wise application of compute_prob to n
plot(n, prob)

```

Code: Computing birthday problem probabilities with sapply

```{r}

# function for computing exact probability of shared birthdays for any n
exact_prob <- function(n){
    prob_unique <- seq(365, 365-n+1)/365   # vector of fractions for mult. rule
    1 - prod(prob_unique)    # calculate prob of no shared birthdays and subtract from 1
}

# applying function element-wise to vector of n values
eprob <- sapply(n, exact_prob)

# plotting Monte Carlo results and exact probabilities on same graph
plot(n, prob)    # plot Monte Carlo results
lines(n, eprob, col = "red")    # add line for exact prob

```


Code: Estimating a practical value of B
This code runs Monte Carlo simulations to estimate the probability of shared birthdays using several B values and plots the results. When B is large enough that the estimated probability stays stable, then we have selected a useful value of B.

```{r}


B <- 10^seq(1, 5, len = 100)    # defines vector of many B values
compute_prob <- function(B, n = 22){    # function to run Monte Carlo simulation with each B
	same_day <- replicate(B, {
    	bdays <- sample(1:365, n, replace = TRUE)
        any(duplicated(bdays))
    })
    mean(same_day)
}

prob <- sapply(B, compute_prob)    # apply compute_prob to many values of B
plot(log10(B), prob, type = "l")    # plot a line graph of estimates 
```

Code: Monte Carlo simulation of stick strategy
```{r}
B <- 10000
stick <- replicate(B, {
	doors <- as.character(1:3)
	prize <- sample(c("car","goat","goat"))    # puts prizes in random order
	prize_door <- doors[prize == "car"]    # note which door has prize
	my_pick  <- sample(doors, 1)    # note which door is chosen
	show <- sample(doors[!doors %in% c(my_pick, prize_door)],1)    # open door with no prize that isn't chosen
	stick <- my_pick    # stick with original door
	stick == prize_door    # test whether the original door has the prize
})
mean(stick)    # probability of choosing prize door when sticking
```

Code: Monte Carlo simulation of switch strategy

```{r}
switch <- replicate(B, {
	doors <- as.character(1:3)
	prize <- sample(c("car","goat","goat"))    # puts prizes in random order
	prize_door <- doors[prize == "car"]    # note which door has prize
	my_pick  <- sample(doors, 1)    # note which door is chosen first
	show <- sample(doors[!doors %in% c(my_pick, prize_door)], 1)    # open door with no prize that isn't chosen
	switch <- doors[!doors%in%c(my_pick, show)]    # switch to the door that wasn't chosen first or opened
	switch == prize_door    # test whether the switched door has the prize
})
mean(switch)    # probability of choosing prize door when switching

```

