---
title: "Assessment: Distributions"
output: html_notebook
---

Exercise 1. Proportions
Histograms and density plots provide excellent summaries of a distribution. But can we summarize even further? We often see the average and standard deviation used as summary statistics: a two number summary! To understand what these summaries are and why they are so widely used, we need to understand the normal distribution.

The normal distribution, also known as the bell curve and as the Gaussian distribution, is one of the most famous mathematical concepts in history. A reason for this is that approximately normal distributions occur in many situations. Examples include gambling winnings, heights, weights, blood pressure, standardized test scores, and experimental measurement errors. Often data visualization is needed to confirm that our data follows a normal distribution.

Here we focus on how the normal distribution helps us summarize data and can be useful in practice.

One way the normal distribution is useful is that it can be used to approximate the distribution of a list of numbers without having access to the entire list. We will demonstrate this with the heights dataset.

Load the height data set and create a vector x with just the male heights:

library(dslabs)
data(heights)
x <- heights$height[heights$sex == "Male"]
Instructions


```{r}
library(dslabs)
data(heights)
x <- heights$height[heights$sex == "Male"]
mean(x > 69 & x <= 72)
```

Exercise 2. Averages and Standard Deviations
Suppose all you know about the height data from the previous exercise is the average and the standard deviation and that its distribution is approximated by the normal distribution. We can compute the average and standard deviation like this:

library(dslabs)
data(heights)
x <- heights$height[heights$sex=="Male"]
avg <- mean(x)
stdev <- sd(x)
Suppose you only have avg and stdev below, but no access to x, can you approximate the proportion of the data that is between 69 and 72 inches?

Given a normal distribution with a mean mu and standard deviation sigma, you can calculate the proportion of observations less than or equal to a certain value with pnorm(value, mu, sigma). Notice that this is the CDF for the normal distribution. We will learn much more about pnorm later in the course series, but you can also learn more now with ?pnorm.

```{r}
library(dslabs)
data(heights)
x <- heights$height[heights$sex=="Male"]
avg <- mean(x)
stdev <- sd(x)
pnorm(72, avg, stdev) - pnorm(69, avg, stdev)
```

Exercise 3. Approximations
Notice that the approximation calculated in the second question is very close to the exact calculation in the first question. The normal distribution was a useful approximation for this case.

However, the approximation is not always useful. An example is for the more extreme values, often called the "tails" of the distribution. Let's look at an example. We can compute the proportion of heights between 79 and 81.

library(dslabs)
data(heights)
x <- heights$height[heights$sex == "Male"]
mean(x > 79 & x <= 81)

```{r}
library(dslabs)
data(heights)
x <- heights$height[heights$sex == "Male"]
avg <- mean(x)
stdev <- sd(x)
exact <- mean(x>79 & x<=81)
approx <- pnorm(81, avg, stdev) - pnorm(79, avg, stdev)
exact/approx
```
Exercise 4. Seven footers and the NBA
Someone asks you what percent of seven footers are in the National Basketball Association (NBA). Can you provide an estimate? Let's try using the normal approximation to answer this question.

Given a normal distribution with a mean mu and standard deviation sigma, you can calculate the proportion of observations less than or equal to a certain value with pnorm(value, mu, sigma). Notice that this is the CDF for the normal distribution. We will learn much more about pnorm later in the course series, but you can also learn more now with ?pnorm.

First, we will estimate the proportion of adult men that are taller than 7 feet.

Assume that the distribution of adult men in the world as normally distributed with an average of 69 inches and a standard deviation of 3 inches.

```{r}
# use pnorm to calculate the proportion over 7 feet (7*12 inches)
1 - pnorm(7*12, 69, 3)
```

Exercise 5. Estimating the number seven footers
Now we have an approximation for the proportion, call it p, of men that are 7 feet tall or taller.

We know that there are about 1 billion men between the ages of 18 and 40 in the world, the age range for the NBA.

Can we use the normal distribution to estimate how many of these 1 billion men are at least seven feet tall?

```{r}
p <- 1 - pnorm(7*12, 69, 3)
round(p * 10^9)
```

Exercise 6. How many seven footers are in the NBA?
There are about 10 National Basketball Association (NBA) players that are 7 feet tall or higher.

```{r}
p <- 1 - pnorm(7*12, 69, 3)
N <- round(p * 10^9)
10/N
```

Exercise 7. Lebron James' height
In the previous exerceise we estimated the proportion of seven footers in the NBA using this simple code:

p <- 1 - pnorm(7*12, 69, 3)
N <- round(p * 10^9)
10/N
Repeat the calculations performed in the previous question for Lebron James' height: 6 feet 8 inches. There are about 150 players, instead of 10, that are at least that tall in the NBA.

```{r}
p <- 1 - pnorm(6*12 + 8, 69, 3)
N <- round(p * 10^9)
150/N
```

