---
title: "Assessment: Summarizing with dplyr"
output: html_notebook
---

Exercise 1. Blood pressure 1
Let's explore the NHANES data. We will be exploring blood pressure in this dataset.

First let's select a group to set the standard. We will use 20-29 year old females. Note that the category is coded with 20-29, with a space in front of the 20! The AgeDecade is a categorical variable with these ages.

To know if someone is female, you can look at the Gender variable.

```{r}
library(dplyr)
library(NHANES)
data(NHANES)
tab <- NHANES %>% filter(AgeDecade == " 20-29" & Gender == "female")
```
Exercise 2. Blood pressure 2
Now we will compute the average and standard deviation for the subgroup we defined in the previous exercise (20-29 year old females), which we will use reference for what is typical.

You will determine the average and standard deviation of systolic blood pressure, which are stored in the BPSysAve variable in the NHANES dataset.

```{r}
library(dplyr)
library(NHANES)
data(NHANES)
ref <- NHANES %>%
  filter(AgeDecade == " 20-29" & Gender == "female") %>%
  summarize(average = mean(BPSysAve, na.rm = TRUE), 
            standard_deviation = sd(BPSysAve, na.rm = TRUE))
ref
```
Exercise 3. Summarizing averages
Now we will repeat the exercise and generate only the average blood pressure for 20-29 year old females. For this exercise, you should review how to use the place holder . in dplyr or the pull function.

```{r}
library(dplyr)
library(NHANES)
data(NHANES)
ref_avg <- NHANES %>%
      filter(AgeDecade == " 20-29" & Gender == "female") %>%
      summarize(average = mean(BPSysAve, na.rm = TRUE)) %>%
      .$average
```

Exercise 4. Min and max
Let's continue practicing by calculating two other data summaries: the minimum and the maximum.

Again we will do it for the BPSysAve variable and the group of 20-29 year old females.

```{r}
library(NHANES)
data(NHANES)
NHANES %>%
      filter(AgeDecade == " 20-29"  & Gender == "female") %>%
      summarize(minbp = min(BPSysAve, na.rm = TRUE), 
                maxbp = max(BPSysAve, na.rm = TRUE))
```

Exercise 5. group_by
Now let's practice using the group_by function.

What we are about to do is a very common operation in data science: you will split a data table into groups and then compute summary statistics for each group.

We will compute the average and standard deviation of systolic blood pressure for females for each age group separately. Remember that the age groups are contained in AgeDecade.

```{r}
library(dplyr)
library(NHANES)
data(NHANES)
NHANES %>%
      filter(Gender == "female") %>%
      group_by(AgeDecade) %>%
      summarize(average = mean(BPSysAve, na.rm = TRUE), 
                standard_deviation = sd(BPSysAve, na.rm=TRUE))
```
Exercise 6. group_by example 2
Now let's practice using group_by some more. We are going to repeat the previous exercise of calculating the average and standard deviation of systolic blood pressure, but for males instead of females.

This time we will not provide much sample code. You are on your own!

```{r}
library(NHANES)
data(NHANES)
NHANES %>%
      filter(Gender == "male") %>%
      group_by(AgeDecade) %>%
      summarize(average = mean(BPSysAve, na.rm = TRUE), 
                standard_deviation = sd(BPSysAve, na.rm=TRUE))
```
Exercise 7. group_by example 3
We can actually combine both of these summaries into a single line of code. This is because group_by permits us to group by more than one variable.

We can use group_by(AgeDecade, Gender) to group by both age decades and gender.

```{r}
library(NHANES)
data(NHANES)
NHANES %>%
      group_by(AgeDecade, Gender) %>%
      summarize(average = mean(BPSysAve, na.rm = TRUE), 
                standard_deviation = sd(BPSysAve, na.rm=TRUE))
```
Exercise 8. Arrange
Now we are going to explore differences in systolic blood pressure across races, as reported in the Race1 variable.

We will learn to use the arrange function to order the outcome acording to one variable.

Note that this function can be used to order any table by a given outcome. Here is an example that arranges by systolic blood pressure.

```{r}
library(dplyr)
library(NHANES)
data(NHANES)
NHANES %>%
      filter(Gender == "male" & AgeDecade==" 40-49") %>%
      group_by(Race1) %>%
      summarize(average = mean(BPSysAve, na.rm = TRUE), 
                standard_deviation = sd(BPSysAve, na.rm=TRUE)) %>%
      arrange(average)
```

