---
title: "Titanic Survival Exercises"
output: html_notebook
---
Put all your new skills together to perform exploratory data analysis on a classic machine learning dataset: Titanic survival!

Background
The Titanic was a British ocean liner that struck an iceberg and sunk on its maiden voyage in 1912 from the United Kingdom to New York. More than 1,500 of the estimated 2,224 passengers and crew died in the accident, making this one of the largest maritime disasters ever outside of war. The ship carried a wide range of passengers of all ages and both genders, from luxury travelers in first-class to immigrants in the lower classes. However, not all passengers were equally likely to survive the accident. We use real data about a selection of 891 passengers to learn who was on the Titanic and which passengers were more likely to survive.

Libraries, Options, and Data 
Be sure that you have installed the titanic package before proceeding.

Define the titanic dataset starting from the titanic library with the following code:

```{r}
options(digits = 3)    # report 3 significant digits
library(tidyverse)
library(titanic)

titanic <- titanic_train %>%
    select(Survived, Pclass, Sex, Age, SibSp, Parch, Fare) %>%
    mutate(Survived = factor(Survived),
           Pclass = factor(Pclass),
           Sex = factor(Sex))
```

Question 1: Variable Types
0.0/3.0 points (graded)
Inspect the data and also use ?titanic_train to learn more about the variables in the dataset. Match these variables from the dataset to their variable type. There is at least one variable of each type (ordinal categorical, non-ordinal categorical, continuous, discrete).

```{r}

```

Question 2: Demographics of Titanic Passengers
0.0/3.5 points (graded)
Make density plots of age grouped by sex. Try experimenting with combinations of faceting, alpha blending, stacking and using variable counts on the y-axis to answer the following questions. Some questions may be easier to answer with different versions of the density plot.

Which of the following are true?
Select all correct answers.

Answer Code and Explanation

A faceted plot is useful for comparing the distributions of males and females for A. Each sex has the same general shape with two modes at the same locations, though proportions differ slightly across ages and there are more males than females.

```{r}
 titanic %>%
ggplot(aes(Age, fill = Sex)) +
geom_density(alpha = 0.2) +
facet_grid(Sex ~ .)

```
A stacked density plot with count on the y-axis is useful for answering B, C and D. The main mode is around age 25 and a second smaller mode is around age 4-5. There are more males than females as indicated by a higher total area and higher counts at almost all ages. With count on the y-axis, it is clear that more males than females are age 40.

```{r}
titanic %>%
    ggplot(aes(Age, y = ..count.., fill = Sex)) +
    geom_density(alpha = 0.2, position = "stack")
        
```
A plot filled by sex with alpha blending helps reveal the answers to E, F and G. There is a higher proportion of females than males below age 17, a higher proportion of males than females for ages 18-35, approximately the same proportion of males and females age 35-55, and a higher proportion of males over age 55. The oldest individuals are male.

```{r}
titanic %>%
    ggplot(aes(Age, fill = Sex)) +
    geom_density(alpha = 0.2)
```
Question 3: QQ-plot of Age Distribution
1 point possible (graded)
Use geom_qq() to make a QQ-plot of passenger age and add an identity line with geom_abline(). Filter out any individuals with an age of NA first. Use the following object as the dparams argument in geom_qq():

```{r}
params <- titanic %>%
    filter(!is.na(Age)) %>%
    summarize(mean = mean(Age), sd = sd(Age))

```

Answer code:

```{r}
titanic %>%
    filter(!is.na(Age)) %>%
    ggplot(aes(sample = Age)) +
    geom_qq(dparams = params) +
    geom_abline()
```

Question 4: Survival by Sex
2/2 points (graded)
To answer the following questions, make barplots of the Survived and Sex variables using geom_bar(). Try plotting one variable and filling by the other variable. You may want to try the default plot, then try adding position = position_dodge() to geom_bar() to make separate bars for each group.

You can read more about making barplots in the textbook section on ggplot2 geometries.

Explanation and Answer Code

A and B can be clearly seen in the barplot of survival status filled by sex. The count of survivors is lower than the count of non-survivors. The bar of survivors is more than half filled by females. Alternatively, the bars can be split by sex with position_dodge, showing the "Female, Survived" bar has a greater height than the "Male, survived" bar. C and D are more clearly seen in the barplot of sex filled by survival status, though they can also be determined from the first barplot. Most males did not survive, but most females did survive.


```{r}
#plot 1 - survival filled by sex
titanic %>%
    ggplot(aes(Survived, fill = Sex)) +
    geom_bar()
# plot 2 - survival filled by sex with position_dodge
titanic %>%
    ggplot(aes(Survived, fill = Sex)) +
    geom_bar(position = position_dodge())
#plot 3 - sex filled by survival
titanic %>%
    ggplot(aes(Sex, fill = Survived)) +
    geom_bar()
```
Question 5: Survival by Age
3/3 points (graded)
Make a density plot of age filled by survival status. Change the y-axis to count and set alpha = 0.2.

Which age group is the only group more likely to survive than die?

```{r}
 titanic %>%
ggplot(aes(Age, y = ..count.., fill = Survived)) +
geom_density(alpha = 0.2)
```
Question 6: Survival by Fare
2.5/2.5 points (graded)
Filter the data to remove individuals who paid a fare of 0. Make a boxplot of fare grouped by survival status. Try a log2 transformation of fares. Add the data points with jitter and alpha blending.

Which of the following are true?

```{r}
          titanic %>%
    filter(Fare > 0) %>%
    ggplot(aes(Survived, Fare)) +
    geom_boxplot() +
    scale_y_continuous(trans = "log2") +
    geom_jitter(alpha = 0.2)
        
```
Question 7: Survival by Passenger Class
0.0/3.0 points (graded)
The Pclass variable corresponds to the passenger class. Make three barplots. For the first, make a basic barplot of passenger class filled by survival. For the second, make the same barplot but use the argument position = position_fill() to show relative proportions in each group instead of counts. For the third, make a barplot of survival filled by passenger class using position = position_fill().

You can read more about making barplots in the textbook section on ggplot2 geometries External link.

Which of the following are true?
Select all correct answers.


Answer code:

```{r}
   # barplot of passenger class filled by survival
titanic %>%
    ggplot(aes(Pclass, fill = Survived)) +
    geom_bar() +
    ylab("Proportion")
# barplot of passenger class filled by survival with position_fill
titanic %>%
    ggplot(aes(Pclass, fill = Survived)) +
    geom_bar(position = position_fill()) +
    ylab("Proportion")
# Barplot of survival filled by passenger class with position_fill
titanic %>%
    ggplot(aes(Survived, fill = Pclass)) +
    geom_bar(position = position_fill()) +
    ylab("Proportion")
```
Question 8: Survival by Age, Sex and Passenger Class
0.0/2.5 points (graded)
Create a grid of density plots for age, filled by survival status, with count on the y-axis, faceted by sex and passenger class.

Which of the following are true?

```{r}
titanic %>%
    ggplot(aes(Age, y = ..count.., fill = Survived)) +
    geom_density(position = "stack") +
    facet_grid(Sex ~ Pclass)
```

